Explore and Summarize Data by Ashish
========================================================
## Dataset Overview:

This tidy data set contains 1,599 red wines with 13 variables on the chemical 
properties of the wine. At least 3 wine experts rated the quality of each wine, 
providing a rating between 0 (very bad) and 10 (very excellent).

### My Point Of View of Analysis:

Which chemical properties influence the quality of red wines?

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(reshape2)
library(purrr)
library(tidyr)
library(GGally)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
red_wine <- read.csv('wineQualityReds.csv', sep = ',')
```

# Univariate Plots Section:

All the Variables in a glance,

```{r echo=FALSE, all_variables}
names(red_wine)
```

### Size of the dataset:

```{r echo=FALSE, Size}
str(red_wine)
```

This dataset(red_wine) tells us about the quality of Red Wine having 1599 
observations of 13 variables.

```{r echo=FALSE, Summary}
summary(red_wine)
```

'quality' variable is discreet and other variables are continuous in the dataset.
'X' looks like some unique number.
'residual.sugar', 'chlorides', 'sulfur.dioxide' are having outliers.

```{r echo=FALSE, delete_X_Variable}
red_wine$X <- NULL
```

Removed 'X' from dataframe as it was just a serial number.

### Histogram of all the variables,

```{r echo=FALSE,message=FALSE, warning=FALSE, All_Variables_Plot}
ggplot(data = melt(red_wine), mapping = aes(x = value)) + 
    geom_histogram(bins = 40) + facet_wrap(~variable, scales = 'free_x')
```

### Quality

```{r echo=FALSE,message=FALSE, warning=FALSE, Quality_1}
ggplot(aes(x = factor(quality)), data = red_wine) +
  geom_bar() + xlab('Quality')
```
```{r echo=FALSE, Quality_2}
summary(red_wine$quality)
```

It's a Normal Distribution.
Dataset is largely concentrated around 5 & 6.

### Fixed Acidity

```{r echo=FALSE,message=FALSE, warning=FALSE, fixed.acidity_1}
ggplot(aes(x = fixed.acidity), data = red_wine) +
  geom_histogram() +
  scale_x_continuous(limits = c(4, 15.90), breaks = seq(4, 15.90, 1)) +
  xlab('Fixed Acidity')
```


```{r echo=FALSE, fixed.acidity_2}
summary(red_wine$fixed.acidity)
```

It's a Right-Skewed Distribution.
Dataset is largely concentrated around 7 & 8.
There are some outliers also.

### Volatile Acidity

```{r echo=FALSE,message=FALSE, warning=FALSE, volatile.acidity_1}
ggplot(aes(x = volatile.acidity), data = red_wine) +
  geom_histogram() +
  scale_x_continuous(limits = c(0.1, 1.58), breaks = seq(0.1, 1.58, 0.1)) +
  xlab('Volatile Acidity')
```
```{r echo=FALSE, volatile.acidity_2}
summary(red_wine$volatile.acidity)
```

Cannot figure out distribution whether it's Bimodal or Normal.

### Citric Acid

```{r echo=FALSE,message=FALSE, warning=FALSE, citric.acid_1}
ggplot(aes(x = citric.acid), data = red_wine) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
  xlab('Citric Acid')
```

```{r echo=FALSE, citric.acid_2}
summary(red_wine$citric.acid)
```

Distribution of 'Citric Acid' is not Normal.

### Residual Sugar

```{r echo=FALSE,message=FALSE, warning=FALSE, residual.sugar_1}
ggplot(aes(x = residual.sugar), data = red_wine) +
  geom_histogram() +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, 1)) +
  xlab('Residual Sugar')
```


```{r echo=FALSE, residual.sugar_2}
summary(red_wine$residual.sugar)
```

log10 Distribution:

```{r echo=FALSE,message=FALSE, warning=FALSE, residual.sugar_3}
ggplot(aes(x = residual.sugar), data = red_wine) +
  geom_histogram() +
  xlab('Residual Sugar') +
  scale_x_log10(breaks = seq(0, 15, 1))
```

Distribution is Right Skewed.
There were some outliers.

### Chlorides

```{r echo=FALSE,message=FALSE, warning=FALSE, chlorides_1}
ggplot(aes(x = chlorides), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('Chlorides')
```

```{r echo=FALSE, chlorides_2}
summary(red_wine$chlorides)
```

log10 Distribution:

```{r echo=FALSE,message=FALSE, warning=FALSE, chlorides_3}
ggplot(aes(x = chlorides), data = red_wine) +
  geom_histogram() +
  scale_x_log10(breaks = seq(0, 1, 0.1)) +
  xlab('log10 Chlorides')
```

Its a Normal Distribution.

### free.sulfur.dioxide

```{r echo=FALSE,message=FALSE, warning=FALSE, free.sulfur.dioxide_1}
ggplot(aes(x = free.sulfur.dioxide), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('Free Sulfur Dioxide')
```
```{r echo=FALSE, free.sulfur.dioxide_2}
summary(red_wine$free.sulfur.dioxide)
```

log10 Distribution:


```{r echo=FALSE,message=FALSE, warning=FALSE, free.sulfur.dioxide_3}
ggplot(aes(x = free.sulfur.dioxide), data = red_wine) +
  geom_histogram() +
  scale_x_log10(breaks = seq(0, 72, 10)) +
  xlab('log10 Free Sulfur Dioxide')
```

It's not Normal.

### Total Sulfur Dioxide

```{r echo=FALSE,message=FALSE, warning=FALSE, total.sulfur.dioxide_1}
ggplot(aes(x = total.sulfur.dioxide), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('Total Sulfur Dioxide')
```

```{r echo=FALSE, total.sulfur.dioxide_2}
summary(red_wine$total.sulfur.dioxide)
```

log10 Distribution:

```{r echo=FALSE,message=FALSE, warning=FALSE, total.sulfur.dioxide_3}
ggplot(aes(x = total.sulfur.dioxide), data = red_wine) +
  geom_histogram() +
  scale_x_log10(breaks = seq(6, 289, 40)) +
  xlab('log10 Total Sulfur Dioxide')
```

It is a Normal Distribution curve.

### density

```{r echo=FALSE,message=FALSE, warning=FALSE, density_1}
ggplot(aes(x = density), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('Density')
```
```{r echo=FALSE, density_2}
summary(red_wine$density)
```

It is a Normal Distribution curve.

### pH
```{r echo=FALSE,message=FALSE, warning=FALSE, pH_1}
ggplot(aes(x = pH), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('pH')
```
```{r echo=FALSE, pH_2}
summary(red_wine$pH)
```

It is a Normal Distribution curve.

### Sulphates

```{r echo=FALSE,message=FALSE, warning=FALSE, sulphates_1}
ggplot(aes(x = sulphates), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('Sulphates')
```


```{r echo=FALSE, sulphates_2}
summary(red_wine$sulphates)
```

log10 Distribution:

```{r echo=FALSE,message=FALSE, warning=FALSE, sulphates_3}
ggplot(aes(x = sulphates), data = red_wine) +
  geom_histogram() +
  scale_x_log10(breaks = seq(0, 2, 0.1)) +
  xlab('log10 Sulphates')
```

It is a Normal Distribution curve.

### alcohol

```{r echo=FALSE,message=FALSE, warning=FALSE, alcohol_1}
ggplot(aes(x = alcohol), data = red_wine) +
  geom_histogram() +
  scale_x_continuous() +
  xlab('Alcohol')
```
```{r echo=FALSE, alcohol_2}
summary(red_wine$alcohol)
```

It's a left-skewed distribution.

### Creating new variable 'quality.rating':

```{r echo=FALSE, quality.rating}
red_wine$quality.rating <- ifelse(red_wine$quality < 5, 'Bad', 
                        ifelse(red_wine$quality < 7, 'Average', 'Excellent'))
red_wine$quality.rating <- ordered(red_wine$quality.rating, levels = 
                                     c('Bad', 'Average', 'Excellent'))
```

### Quality Rating:

```{r echo=FALSE,message=FALSE, warning=FALSE, quality.rating_1}
ggplot(aes(x = quality.rating), data = red_wine) +
  geom_bar() +
  xlab('Quality Rating')
```

```{r echo=FALSE, quality.rating_2}
summary(red_wine$quality.rating)
```

# Univariate Analysis:

### What is the structure of your dataset?

There were 1599 obs. of  13 variables.

### What is/are the main feature(s) of interest in your dataset?

The main interest was 'quality' as it is the key factor of wine's taste.

### What other features in the dataset do you think will help support your
### investigation into your feature(s) of interest?

'alcohol' & 'pH' will also play a important ingredient for wines taste.

### Did you create any new variables from existing variables in the dataset?

'quality.rating' was created to show the ratings of the quality.

### Of the features you investigated, were there any unusual distributions?
### Did you perform any operations on the data to tidy, adjust, or change the 
### form of the data? If so, why did you do this?

I removed variable 'X' from dataset as it was of no use.

There were many outliers(can be seen above in Box Plots & Histograms). 
To know the actual distribution many variables were log transformed.

# Bivariate Plots Section:

Through 'GGally' package we can see the correlation between the variables:

```{r echo=FALSE,message=FALSE, warning=FALSE, Bivariate_Plots}
ggcorr(red_wine, label = TRUE, 
       label_alpha = TRUE, legend.size = 12, label_size = 4, size = 4, 
       hjust = 0.85, layout.exp = 2)
```

Variable's with strong Positive Correlation:

 * alcohol & quality
 * free sulfur dioxide & total sulfur dioxide
 * density & fixed acidity
 * citric acid & fixed acidity

```{r echo=FALSE,message=FALSE, warning=FALSE, Positive_Correlation}
p1 <- ggplot(aes(x = alcohol, y = quality), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

p2 <- ggplot(aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide), 
      data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

p3 <- ggplot(aes(x = density, y = fixed.acidity), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

p4 <- ggplot(aes(x = fixed.acidity, y = citric.acid), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

library(gridExtra)

grid.arrange(p1, p2, p3, p4, ncol = 2)
  
```

Variable's with strong Negative Correlation:

 * density & alcohol
 * pH & citric acid
 * volatile acidity & fixed acidity
 * pH & fixed acidity
 * volatile acidity & citric acid
 
```{r echo=FALSE,message=FALSE, warning=FALSE, Negative_Correlation}
q1 <- ggplot(aes(x = alcohol, y = density), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

q2 <- ggplot(aes(x = pH, y = citric.acid), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

q3 <- ggplot(aes(x = volatile.acidity, y = fixed.acidity), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

q4 <- ggplot(aes(x = pH, y = fixed.acidity), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

q5 <- ggplot(aes(x = volatile.acidity, y = citric.acid), data = red_wine) +
  geom_smooth(method = 'lm', color = 'Red') + geom_jitter(alpha = (1/10))

library(gridExtra)

grid.arrange(q1, q2, q3, q4, q5, ncol = 2)
  
```

Same thing can be visualised from the above graphs.

Now, we will compare our main intrest i.e 'quality.rating' with alcohol,

```{r echo=FALSE, Factor_quality.rating}
red_wine$quality.rating <- factor(red_wine$quality.rating,
                                  labels = c('Bad', 'Average', 'Excellent'))
```

```{r echo=FALSE, Factor_quality}
red_wine$quality <- factor(red_wine$quality,
                                  labels = c('3', '4', '5', '6', '7', '8'))
```


```{r echo=FALSE, quality.rating_alcohol}
ggplot(aes(x = quality.rating, y = alcohol), data = red_wine) +
  geom_boxplot(fill = "gold1", colour = "Red") + 
  ggtitle("Quality Rating Vs Alcohol") + geom_jitter(alpha = (1/10))
```
  
```{r echo=FALSE, quality_alcohol}
ggplot(aes(x = quality, y = alcohol), data = red_wine) +
  geom_boxplot(fill = "gold1", colour = "Red") + ggtitle("Quality Vs Alcohol") +
  geom_jitter(alpha = (1/10))
``` 

As we have seen in above chart quality and alcohol are having a positive
correlation, which can also be seen from above box plots.

As we can see that 'quality.rating' as Excellent are having alcohol quantity 
more than 'Bad'   and   'Average'.
  
Sample size for 'Average' looks more than other two i.e. 'Excellent' & 'Bad'.

```{r echo=FALSE, quality_volatile.acidity}
ggplot(aes(x = quality, y = volatile.acidity), data = red_wine) +
  geom_boxplot(fill = "gold1", colour = "Red") + 
  ggtitle("Quality Vs Volatile Acidity") + geom_jitter(alpha = (1/10))
``` 

# Bivariate Analysis:

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Our main intrest was with 'quality', which was having a positive correlation
with 'alcohol'. The alcohol which was having grade '7' and '8' were having 
good amount of alcohol.
  
Red Wine having less volatile acidity were having good quality.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Density is strongly correlated with alcohol (negatively).

### What was the strongest relationship you found?

> Positive:

  * fixed.acidity:citirc.acid = 0.7
  * fixed.acidity:density = 0.7
  * free.sulfur.dioxide:total.sulfur.dioxide = 0.7
  
> Negative:

  * fixed.acidity:pH = 0.7

# Multivariate Plots Section:

I will be plotting alcohol, density, sulaphates, quality and volatile acidity
  as they were strongly related with 'quality' of red wine.

```{r echo=FALSE, Multivariate_Plots_1}
ggplot(aes(x = volatile.acidity , y = sulphates, color = factor(quality)), data = red_wine) +
  geom_point() + scale_colour_brewer(palette = "RdYlGn")
```

As we can see from the plot that red wine with quality '8' is around left most
part of the graph. Through this we can incur that red wine with less volatile
acidity are good in quality.

```{r echo=FALSE, Multivariate_Plots_2}
ggplot(aes(x = density , y = sulphates, color = factor(quality)), data = red_wine) +
  geom_point() + scale_colour_brewer(palette = "RdYlGn")
```

Same goes with 'density' also. Wines having less density are having good
quality.
  
### Models:


```{r echo=FALSE, message=TRUE}
m1 <- lm(formula = as.numeric(quality) ~ alcohol, data = red_wine)
summary(m1)
```

```{r echo=FALSE, message=TRUE}
m2 <- update(m1, . ~ . + sulphates)
summary(m2)
```
```{r echo=FALSE, message=TRUE}
m3 <- update(m2, . ~ . + volatile.acidity)
summary(m3)
```

# Multivariate Analysis:

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Since 'quality' was highly correlated with density, sulaphates, alcohol and 
volatile acidity.So I plotted sulaphates, alcohol, volatile acidity and 
quality (In Multivariate Plot 1). So we can see that with increase in volatile
acid quality is decreasing.
  
In  Multivariate Plot 2,
I plotted sulaphates, alcohol, density and quality. 
So here we can see that with increase in density quality is decreasing.

### Were there any interesting or surprising interactions between features?

Here I did not find any thing unusual or interesting may be I dont have domain
knowledge to check those variables which may impact analysis. It was purely
based on the correlations.

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

Yes, I created three models. R-squared values were very less for all the three
models(Higher is better with 1 being the best),which were not as much convincing
to declare as a linear model for this dataset.

Here p-value indicates there is a significant relationship described by the 
model because the values are less than 0.05.So we reject the null hypothesis.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE,message=FALSE, warning=FALSE, Plot_One_1}
ggplot(aes(x = factor(quality)), data = red_wine) +
  geom_bar() + xlab('Quality') + ylab('Number of Red Wines')
```
```{r echo=FALSE, message=TRUE}
summary(red_wine$quality)
```

### Description One

The distribution of red wine quality appears to be normal. Although the rating
scale is between 0 and 10, there exists no wine that is rated 1, 2, 9 or 10.

### Plot Two

```{r echo=FALSE, Plot_Two}
ggplot(aes(x = factor(quality), y = volatile.acidity), data = red_wine) +
  geom_boxplot(fill = "gold1", colour = "Red") + ggtitle("Quality Vs Volatile Acidity") +
  geom_jitter(alpha = (1/10)) + ylab("Volatile Acidity (g / dm^3)")
```

### Description Two

Volatile Acid is not having a very strong correlation with quality. But it was
significant enough.
Red Wine having less volatile acidity were having good quality.

### Plot Three
```{r echo=FALSE, Plot_Three}
r1 <- ggplot(aes(x = volatile.acidity , y = sulphates, color = factor(quality)), 
      data = red_wine) + geom_point() + scale_colour_brewer(palette = "RdYlGn")+
      ggtitle('Relation Between Volatile Acidity, Sulphates, Quality') +
      xlab("Volatile Acidity (g / dm^3)") + ylab("Sulphates(g / dm3)")

r2 <- ggplot(aes(x = density , y = sulphates, color = factor(quality)),
      data = red_wine) + geom_point() + scale_colour_brewer(palette = "RdYlGn")+
      ggtitle('Relation Between Density, Sulphates, Quality') +
      xlab("Density (g / cm^3)") + ylab("Sulphates(g / dm3)")

library(gridExtra)

grid.arrange(r1, r2, nrow = 2)
```

### Description Three

'quality' is highly correlated with density, sulaphates, alcohol and volatile
acidity. With the increase in density and volatile acidity, the quality of
Red Wine increases.

------

# Reflection

Red Wine data set contains information on 1599 wines across 13 variables.
Then exploratory data analysis was done in which at first a univariate, then 
bivariate and finally multivariate examinations were performed to know 
relations between its features.

And I have the feeling that some interesting relations(Multivariate) may have 
been left aside among the many permutations of variables in the dataset.
I think good domain knowledge about red wine would have helped a lot while doing
analysis of this dataset. As I could have known relation between variables which
impacts the quality. This would have certainly improve the accuracy of the 
prediction models.

Another challenge was factoring the quality variable, which was rectified by 
reviewer how to efficiently with ordinal variables.
